# Philosophy of PPBstats {#philo}

## What is Participatory Plant Bredding ?

- **decentralized the selection**

When decentralizing the selection in target environment [@desclaux_changes_2008], the phenotypic trait is the sum of genetic ($G$), environmental ($E$) and interaction between the two ($G \times E$) : $P = G + E + G \times E$.
Broad sens heritability of a given trait is then: 

$h^2_{sl} = \frac{var(G)}{var(G) + var(E) + var(G\times E)}$

Heritability taking into account the network is therefore inferior or equal to heritability on a given site because $var(G\times E) > 0$.
When facing a wide diversity of agroecological environment and practices, decentralized breeding is a key point to select adapted varieties to local agro-systems.

- **involve all actors in the breeding decision process**: farmers, technicians, researchers, facilitators, consumers ... Such involvements empower all actors and answer the real needs of actors  [@sperling_framework_2001].


## Objectives of PPBstats

PPBstats aims to perform analysis found within PPB programmes at four levels:

- network of seeds circulation (Section \@ref{network))

- agronomic trials (Section \@ref(agronomic) in order to set, describe and analyse balanced and unbalanced trials in decentralized participatory plant breeding programmes.
The statistical procedures are based on frequentist and bayesian approaches.

- organoleptic tests (Section \@ref(organoleptic))

- molecular experiments (Section \@ref(molecular))


Up to now, only agronomic analysis is under development. 
Please, feel free to contribute to other aspects of analysis, see(here)[https://github.com/priviere/PPBstats/releases/download/v0.22/contribute_PPBstats.pdf] for more information.


## Decision tree to choose the right analysis according to question and experimental design


!!! 
TO DO : integrer decision tree avec même normencalture que dans la vignette + référer aux parties adéquates

+ attente complément Camille

!!!

## Data formating 

To be analyzed by PPBstats, the data must have a certain format.
Up to now, there are two formats supported:

- `data-network`
- `data-agro`

To check and give the right format, you must use the `format_data_PPBstats` function.
Each format are detailed in their corresponding section.


## Let's go!

To continue, load the current development version of the package (here)[https://github.com/priviere/PPBstats/archive/v0.22.zip] 

unzip the folder and then run 

```{r, eval=FALSE}
devtools::install("PPBstats-0.22")
```

Note that `JAGS` must be installed on your computer. See http://mcmc-jags.sourceforge.net/ for more details.

Then you can load the package :

```{r,warning=FALSE,message=FALSE}
library(PPBstats)
```

and download from internet the data used in this vignette (this is useful to earn lots of time!) [here](https://github.com/priviere/PPBstats/releases/download/v0.20/data_PPBstats.zip).

The following examples have been performed with the following R version :
```{r}
unlist(sessionInfo()$R.version)
```


