# Philosophy of PPBstats {#philo}

## What is Participatory Plant Bredding ?

- **decentralize the selection**

When decentralizing the selection in target environment [@desclaux_changes_2008], the phenotypic trait is the sum of genetic ($G$), environmental ($E$) and interaction between the two ($G \times E$) : $P = G + E + G \times E$.
Broad sens heritability of a given trait is then: 

$h^2_{sl} = \frac{var(G)}{var(G) + var(E) + var(G\times E)}$

Heritability taking into account the network is therefore inferior or equal to heritability on a given site because $var(G\times E) > 0$.
When facing a wide diversity of agroecological environment and practices, decentralized breeding is a key point to select adapted varieties to local agro-systems.

- **involve all actors in the breeding decision process**: farmers, technicians, researchers, facilitators, consumers ... Such involvements empower all actors and answer the real needs of actors  [@sperling_framework_2001].


## Objectives of `PPBstats`

`PPBstats` aims to perform analysis found within PPB programmes at four levels:

- network of seeds circulation (section \@ref(network))

- agronomic trials (section \@ref(agronomic) in order to set, describe and analyse balanced and unbalanced trials in decentralized participatory plant breeding programmes.
The statistical procedures are based on frequentist and bayesian approaches.

- organoleptic tests (section \@ref(organoleptic))

- molecular experiments (section \@ref(molecular))

Please, feel free to contribute to other aspects of analysis, see section \@ref(contributions) for more information.


## Analysis according to the objectives

Five main objectives dealing with PPB programmes can be :

- **Study network of seed circulation** to have insight on how exchanges are structured through network analysis (section \@ref(network)).

- **Improve the prediction of a target variable for selection**: this refers to non parametric and multivariate analysis, regression, classificationand regression trees, random forest.
Only multivariate analyses are proposed in `PPBstats` in section \@ref(multivariate-analysis).

- **Compare different germplasms on each location for selection**.
This can be done through 
     - agronomic analysis within family 1 of analysis (section \@ref(family-1))
         - classic anova (section \@ref(classic-anova))
         - spatial model (section \@ref(spatial-analysis))
         - bayesian hierarchical model (section \@ref(model-1))
     - organoleptic analysis (section \@ref(organoleptic))

- **Study response to selection of germplasm over several locations for selection**.
This can be done through agronomic analysis within family 2 of analysis (section \@ref(family-2)):
    - incomplete block design (section \@ref(ibd))
    - AMMI (section \@ref(ammi))
    - GGE (section \@ref(gge))
    - migrant-residant (section) \@ref(migrant-residant))
    - bayesian model (section \@ref(model-2))

- **Study diversity structure and identify complementary or similar parents for cross**.
This can be done through 
    - agronomic:
        - multivariate analysis and clustering mentionned in section \@ref(multivariate-analysis)
    - molecular analysis (section \@ref(molecular))

Each of these objective is explained throught exemple for each analyse in its corresponding section.

## Data formating 

To be analyzed by `PPBstats`, the data must have a certain format.
Up to now, there are two formats supported:

- `data-network`
- `data-agro`

To check and give the right format, you must use the `format_data_PPBstats()` function.
Each format are detailed in their corresponding section.


## Let's go!

To continue, load the current development version of the package [here](https://github.com/priviere/PPBstats/archive/v0.22.zip)

unzip the folder and then run 

```{r, eval=FALSE}
devtools::install("PPBstats-0.22")
```

Note that `JAGS` must be installed on your computer. See http://mcmc-jags.sourceforge.net/ for more details.

Then you can load the package :

```{r,warning=FALSE,message=FALSE}
library(PPBstats)
```

and download from internet the data used in this vignette (this is useful to earn lots of time!) [here](https://github.com/priviere/PPBstats/releases/download/v0.20/data_PPBstats.zip).

The following examples have been performed with the following R version :
```{r}
unlist(sessionInfo()$R.version)
```


