## Unipart network for seed lots analysis
This section deals with unipart network that represents the relationships between seed lots.

### Steps with `PPBstats`

- Format the data with `format_data_PPBstats()`
- get descriptive plot with `plot()`

### Format the data
The format required is a data frame with the following compulsory columns as factor:

- `"seed_lot_parent"` : name of the parent seed lot in the relationship
- `"seed_lot_child"` ; name of the child seed lots in the relationship
- `"relation_type"` : the type of relationship between the seed lots
- `"relation_year_start"` : the year when the relationship starts
- `"relation_year_end"` : the year when the relationship stops
- `"germplasm_parent"` : the germplasm associated to the parent seed lot
- `"location_parent"` : the location associated to the parent seed lot
- `"year_parent"` : the year of the last relationship of the parent seed lot
- `"germplasm_child"` : the germplasm associated to the child seed lot
- `"location_child"` : the location associated to the child seed lot
- `"year_child"` : represents the year of the last relation event of the child seed lot

Possible options are : `"long_parent"`, `"lat_parent"`, `"long_child"`, `"lat_child"` to get map representation, supplementary variables with tags: `"_parent"`, `"_child"` or `"_relation"`.

The format of the data are checked by the function `format_data_PPBstats()` with the following arguments :

- `type` : `"data_network"`
- `network_part` : `"unipart"`
- `vertex_type` : `"seed_lots"`

The function returns list of `igraph` object^[http://igraph.org/r/] coming from `igraph::graph_from_data_frame()`.

```{r}
data(data_network_unipart_sl)
head(data_network_unipart_sl)
```

```{r}
net_unipart_sl = format_data_PPBstats(
  type = "data_network",
  data = data_network_unipart_sl, 
  network_part = "unipart", 
  vertex_type =  "seed_lots")
length(net_unipart_sl)
head(net_unipart_sl)
```

### Describe the data

The different representations are done with the `plot()` function.

For network representation, set `plot_type = "network"` diffusion event are displayed with a curve.
`in_col` can be settled to customize color of vertex.

```{r, message=FALSE, warning=FALSE}
p_net = plot(net_unipart_sl, plot_type = "network", in_col = "location")
p_net
```

In order to get the network organized in a chronologiical order and by location, set `organize_sl = TRUE`.
This representation is possible if the seed lots are under the following format : `GERMPLASM_LOCATION_YEAR_DIGIT`.

```{r, warning=FALSE}
p_net_org = plot(net_unipart_sl, plot_type = "network", organize_sl = TRUE)
p_net_org
```

To have information on the seed lots that are represented, `plot_type = "barplot"` can be used.
Choose what to represent on the x axis and in color as well as the number of parameter per plot.

```{r, warning=FALSE}
p_bar = plot(net_unipart_sl, plot_type = "barplot", in_col = "location", 
                          x_axis = "germplasm", nb_parameters_per_plot_x_axis = 5, 
                          nb_parameters_per_plot_in_col = 5)
p_bar[[1]]$barplot$`germplasm-1|location-1` # first element of the plot
```

Barplot can also be use to study the relation within the network.
The name of the relation must be put in the argument `vec_variables`.
The results is a list of two elements for each variable:

- nb_received: number of seed lots that end the relation
- nb_given: number of seed lots that start the relation

```{r,warning=FALSE}
p_bar = plot(net_unipart_sl, plot_type = "barplot", vec_variables = "diffusion",
                  nb_parameters_per_plot_x_axis = 100, x_axis = "location", in_col = "year")
p_bar
```

Location present on the network can be displayed on a map with `plot_type = "map"`

```{r,warning=FALSE,message=FALSE}
p_map = plot(net_unipart_sl, plot_type = "map", labels_on = "location")
p_map
```

It can be interesting to plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `variable`:

```{r, warning=FALSE}
nb_values = 30
data_to_pie = data.frame( # = data_agro : ajouter id_seed_lot dans les format agro
  id = rep(c("germ-4_loc-4_2009_0001", "germ-9_loc-4_2009_0001", "germ-10_loc-3_2009_0001", "germ-12_loc-3_2007_0001", "germ-11_loc-2_2009_0001", "germ-10_loc-2_2009_0001"), each = nb_values),
  location = rep(c("loc-1", "loc-1", "loc-3", "loc-3", "loc-2", "loc-2"), each = nb_values),
  year = rep(c("2009", "2008", "2007", "2007", "2009", "2009"), each = nb_values),
  germplasm = rep(c("germ-7", "germ-2", "germ-6", "germ-4", "germ-5", "germ-13"), each = nb_values),
  block = 1,
  X = 1,
  Y = 1,
  y1 = rnorm(nb_values*6, 10, 2), # quanti
  y2 = rep(c("cat1", "cat1", "cat2", "cat3", "cat3", "cat4"), each = nb_values)  # quali
)
```


```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = plot(net_unipart_sl, data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1
```


```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = plot(net_unipart_sl, data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2
```

or on the network with a pie with `plot_type = "network"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r, warning=FALSE}
# y1 is a quantitative variable
p_net_pies_y1 = plot(net_unipart_sl, data_to_pie, plot_type = "network", vec_variables = "y1")
p_net_pies_y1
```

```{r,warning=FALSE}
# y2 is a qualitative variable
p_net_pies_y2 = plot(net_unipart_sl, data_to_pie, plot_type = "network", vec_variables = "y2")
p_net_pies_y2
```

The same can be done regarding relation type of the network.
This can be displayed on a map but not on a network.

```{r,warning=FALSE,message=FALSE}
p_map_pies_diff = plot(net_unipart_sl, plot_type = "map", vec_variables = "diffusion")
p_map_pies_diff
```

Here the pies represent the repartition of the number of seed lots.
