## Describe the data {#describe-data-network}

The different representations are done with the `plot` function.

### unipart network on seed lots

For network representation, set `plot_type = "network"` diffusion event are displayed with a curve.
`in_col` can be settled to customize color of vertex.

```{r, message=FALSE}
p_net = plot(net_unipart_sl, plot_type = "network", in_col = "location")
p_net
```

In order to get the network organized in a chronolical order and by farm, set `organize_sl = TRUE`.
This representation is possible if the seed lots are under the following format : `GERMPLASM_LOCATION_YEAR_DIGIT`.

```{r}
p_net_org = plot(net_unipart_sl, plot_type = "network", in_col = "location", 
                              organize_sl = TRUE)
p_net_org
```

To have information on the seed lots that are represented, `plot_type = "barplot"` can be used.
Choose what to represent on the x axis and in color as well as the number of parameter per plot.

```{r}
p_bar = plot(net_unipart_sl, plot_type = "barplot", in_col = "location", 
                          x_axis = "germplasm", nb_parameters_per_plot_x_axis = 5, 
                          nb_parameters_per_plot_in_col = 5)
p_bar[[1]]$barplot$`germplasm-1|location-1` # first element of the plot
```


### unipart network on location

For network representation, set `plot_type = "network"` diffusion event are display with curve.
`in_col` can be settle to customize color of vertex.


```{r}
p_net = plot(net_unipart_location_g, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per germplasm, the first element with all the data
p_net$C1
```

```{r}
p_net = plot(net_unipart_location_y, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2012-2013-2014`
```


### bipart network on germplasm and location

```{r}
p_net = plot(net_bipart, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2014`
```


With barplots, it represents the number of edges per vertex for each germplasm and each location.
```{r}
p_bar = plot(net_bipart, plot_type = "barplot")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2010-2011-2012-2013-2014`$barplot
p_bar$germplasm
p_bar$location
```

