## Describe the data {#describe-data-network}

The different representations are done with the `plot` function.

### unipart network on seed lots

For network representation, set `plot_type = "network"` diffusion event are displayed with a curve.
`in_col` can be settled to customize color of vertex.

```{r, message=FALSE, warning=FALSE}
p_net = plot(net_unipart_sl, plot_type = "network", in_col = "location")
p_net
```

In order to get the network organized in a chronologiical order and by location, set `organize_sl = TRUE`.
This representation is possible if the seed lots are under the following format : `GERMPLASM_LOCATION_YEAR_DIGIT`.

```{r, warning=FALSE}
p_net_org = plot(net_unipart_sl, plot_type = "network", organize_sl = TRUE)
p_net_org
```

To have information on the seed lots that are represented, `plot_type = "barplot"` can be used.
Choose what to represent on the x axis and in color as well as the number of parameter per plot.

```{r, warning=FALSE}
p_bar = plot(net_unipart_sl, plot_type = "barplot", in_col = "location", 
                          x_axis = "germplasm", nb_parameters_per_plot_x_axis = 5, 
                          nb_parameters_per_plot_in_col = 5)
p_bar[[1]]$barplot$`germplasm-1|location-1` # first element of the plot
```

Barplot can also be use to study the relation within the network.
The name of the relation must be put in the argument `vec_variables`.
The results is a list of two elements for each variable:

- nb_received: number of seed lots that end the relation
- nb_given: number of seed lots that start the relation

```{r,warning=FALSE}
p_bar = plot(net_unipart_sl, plot_type = "barplot", vec_variables = "diffusion",
                  nb_parameters_per_plot_x_axis = 100, x_axis = "location", in_col = "year")
p_bar
```

Location present on the network can be displayed on a map with `plot_type = "map"`

```{r,warning=FALSE,message=FALSE}
p_map = plot(net_unipart_sl, plot_type = "map", labels_on = "location")
p_map
```

It can be interesting to plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `variable`:

```{r, warning=FALSE}
nb_values = 30
data_to_pie = data.frame( # = data_agro : ajouter id_seed_lot dans les format agro
  id = rep(c("germ-4_loc-4_2009_0001", "germ-9_loc-4_2009_0001", "germ-10_loc-3_2009_0001", "germ-12_loc-3_2007_0001", "germ-11_loc-2_2009_0001", "germ-10_loc-2_2009_0001"), each = nb_values),
  location = rep(c("loc-1", "loc-1", "loc-3", "loc-3", "loc-2", "loc-2"), each = nb_values),
  year = rep(c("2009", "2008", "2007", "2007", "2009", "2009"), each = nb_values),
  germplasm = rep(c("germ-7", "germ-2", "germ-6", "germ-4", "germ-5", "germ-13"), each = nb_values),
  block = 1,
  X = 1,
  Y = 1,
  y1 = rnorm(nb_values*6, 10, 2), # quanti
  y2 = rep(c("cat1", "cat1", "cat2", "cat3", "cat3", "cat4"), each = nb_values)  # quali
)
```


```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = plot(net_unipart_sl, data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1
```


```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = plot(net_unipart_sl, data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2
```

or on the network with a pie with `plot_type = "network"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r, warning=FALSE}
# y1 is a quantitative variable
p_net_pies_y1 = plot(net_unipart_sl, data_to_pie, plot_type = "network", vec_variables = "y1")
p_net_pies_y1
```

```{r,warning=FALSE}
# y2 is a qualitative variable
p_net_pies_y2 = plot(net_unipart_sl, data_to_pie, plot_type = "network", vec_variables = "y2")
p_net_pies_y2
```

The same can be done regarding relation type of the network on map but not on network.

```{r,warning=FALSE,message=FALSE}
p_map_pies_diff = plot(net_unipart_sl, plot_type = "map", vec_variables = "diffusion")
p_map_pies_diff
```

Here the pies represent the repartition of the number of seed lots.

### unipart network on location

For network representation, set `plot_type = "network"` diffusion event are display with curve.
`in_col` can be settle to customize color of vertex.
The curve between location represent the diffusion, the number of diffusion is displayed on a color scale.


```{r,warning=FALSE}
p_net = plot(net_unipart_location_g, plot_type = "network", 
                          labels_on = "location", labels_size = 4)
names(p_net) # one element per germplasm, the first element with all the data
p_net$`germ-2`
```

```{r,warning=FALSE}
p_net = plot(net_unipart_location_y, plot_type = "network", 
                          labels_on = "location", labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2007-2008-2009`
```

With barplots, it represents the number of germplasm received or given.
```{r}
p_bar = plot(net_unipart_location_y, plot_type = "barplot", x_axis = "location", in_col = "germplasm")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2007-2008-2009`
p_bar$barplot$received
p_bar$barplot$given
```

Location present on the network can be displayed on a map with `plot_type = "map"`.

```{r,warning=FALSE,message=FALSE}
p_map = PPBstats:::plot.data_network(net_unipart_location_y[1], plot_type = "map", labels_on = "location")
# Note if you want to do it on all element of the list, you should use 
# plot(net_unipart_location_y, plot_type = "map", labels_on = "location")
# Here we use PPBstats:::plot.data_network only not to ask to often the map server that may bug if there are too many query
# We use ::: because the function is not exported as it is an S3 method
p_map$`2007-2008-2009`
```

As well as plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = PPBstats:::plot.data_network(net_unipart_location_y[1], data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1$`2007-2008-2009`
```

```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = PPBstats:::plot.data_network(net_unipart_location_y[1], data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2$`2007-2008-2009`
```

Note that it is not possible to display plot with `plot_type = "network"`.



### bipart network on germplasm and location

```{r,warning=FALSE}
p_net = plot(net_bipart, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2009`
```


With barplots, it represents the number of edges per vertex for each germplasm and each location.
```{r,warning=FALSE}
p_bar = plot(net_bipart, plot_type = "barplot")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2005-2006-2007-2008-2009`$barplot
p_bar$germplasm
p_bar$location
```


Location present on the network can be displayed on a map with `plot_type = "map"`.

```{r,warning=FALSE,message=FALSE}
p_map = PPBstats:::plot.data_network(net_bipart[1], plot_type = "map", labels_on = "location")
p_map$`2005-2006-2007-2008-2009`
```

As well as plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1$`2005-2006-2007-2008-2009`
```

```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2$`2005-2006-2007-2008-2009`
```

Note that it is not possible to display plot with `plot_type = "network"`.


