## Unipart network for location analysis

This section deals with unipart network that represent relationship of germplasm diffusion between locations: a data frame with the following compulsory columns (same as above): `"location_parent"`, `"location_child"`, `"relation_year_start"`, `"relation_year_end"`.
Possible options are : `"germplasm_parent"`, `"year_parent"`, `"germplasm_child"`, `"year_child"`.
Other possibles option are : `"long_parent"`, `"lat_parent"`, `"long_child"`, `"lat_child"` to get map representation.

The format of the data are checked by the function `format_data_PPBstats()`. 

The following arguments can be used :

- `type` : `"data_network"`
- `network_part` : `"unipart"` 
- `vertex_type` : `"location"`
- `network_split` : `"germplasm"` or `"relation_year_start"`.


Possible values of arguments regarding network formats are displayed in Table \@ref(tab:network-format-arg).

| | `network_part` | `vertex_type` | `network_split` | 
| --- | --- | --- | --- | 
| unipart for seed lots format |  `unipart` or  `bipart` |  `seed_lots` or `location` or  `c("germplasm", "location")` |  `NULL` or `germplasm` or `relation_year_start` | 
| unipart for location format |  `unipart` |  `location` |  `germplasm` or `relation_year_start` |
| bipart for germplasm and location format |  `bipart` |  `c("germplasm", "location")` |  `NULL` | 
Table: (\#tab:network-format-arg) Possible values of argument (in colum) regarding network format (in row).

The following sections give examples for each network format.
The function returns list of `igraph` object^[http://igraph.org/r/] coming from `igraph::graph_from_data_frame()`.

### Format data
```{r}
data(data_network_unipart_sl)
head(data_network_unipart_sl)
```
```{r}
net_unipart_location_g = format_data_PPBstats(
  type = "data_network",
  data = data_network_unipart_sl, 
  network_part = "unipart", 
  vertex_type =  "location",
  network_split = "germplasm")
```
For `network_split = "germplasm"`, it returns a list with as many elements as germplam in the data
as well as all germplasms merged in the first element of the list.
```{r}
names(net_unipart_location_g)
```

```{r}
net_unipart_location_y = format_data_PPBstats(
  type = "data_network",
  data = data_network_unipart_sl,
  network_part = "unipart", 
  vertex_type =  "location",
  network_split = "relation_year_start")
```

For `network_split = "relation_year_start"`, it returns a list with as many elements as year in the data
as well as all years merged in the first element of the list.
```{r}
names(net_unipart_location_y)
```

### Describe the data

The different representations are done with the `plot` function.

For network representation, set `plot_type = "network"` diffusion event are display with curve.
`in_col` can be settle to customize color of vertex.
The curve between location represent the diffusion, the number of diffusion is displayed on a color scale.


```{r,warning=FALSE}
p_net = plot(net_unipart_location_g, plot_type = "network", 
                          labels_on = "location", labels_size = 4)
names(p_net) # one element per germplasm, the first element with all the data
p_net$`germ-2`
```

```{r,warning=FALSE}
p_net = plot(net_unipart_location_y, plot_type = "network", 
                          labels_on = "location", labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2007-2008-2009`
```

With barplots, it represents the number of germplasm received or given.
```{r}
p_bar = plot(net_unipart_location_y, plot_type = "barplot", x_axis = "location", in_col = "germplasm")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2007-2008-2009`
p_bar$barplot$received
p_bar$barplot$given
```

Location present on the network can be displayed on a map with `plot_type = "map"`.

```{r,warning=FALSE,message=FALSE}
p_map = PPBstats:::plot.data_network(net_unipart_location_y[1], plot_type = "map", labels_on = "location")
# Note if you want to do it on all element of the list, you should use 
# plot(net_unipart_location_y, plot_type = "map", labels_on = "location")
# Here we use PPBstats:::plot.data_network only not to ask to often the map server that may bug if there are too many query
# We use ::: because the function is not exported as it is an S3 method
p_map$`2007-2008-2009`
```

As well as plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = PPBstats:::plot.data_network(net_unipart_location_y[1], data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1$`2007-2008-2009`
```

```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = PPBstats:::plot.data_network(net_unipart_location_y[1], data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2$`2007-2008-2009`
```

Note that it is not possible to display plot with `plot_type = "network"`.

