## Bipart network analysis
This section deals with bipart network that represents "which location has which germplasm which year".

### Steps with `PPBstats`

- Format the data with `format_data_PPBstats()`
- get descriptive plot with `plot()`

### Format the data

The format required is a data frame with the following compulsory columns as factor: `"germplasm"`, `"location"`, `"year"`.
Possible options are : `"long"`, `"lat"` to get map representation

Note that data frame with unipart network for seed lots format can also be used.

The format of the data are checked by the function `format_data_PPBstats()` with the following arguments :

- `type` : `"data_network"`
- `network_part` : `"bipart"`
- `vertex_type` : `c("germplasm", "location")`


The function returns list of `igraph` object^[http://igraph.org/r/] coming from `igraph::graph_from_data_frame()`.

```{r}
data(data_network_unipart_sl)
head(data_network_unipart_sl)
```

```{r}
net_bipart = format_data_PPBstats(
  type = "data_network",
  data = data_network_unipart_sl, 
  network_part = "bipart", 
  vertex_type =  c("germplasm", "location")
  )
```

For bipart network, it returns a list with as many elements as year in the data 
as well as all years merged in the first element of the list.
If no year are provided into the data, all information are merged.

```{r}
names(net_bipart)
```

### Describe the data

The different representations are done with the `plot()` function.

```{r,warning=FALSE}
p_net = plot(net_bipart, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2009`
```


With barplots, it represents the number of edges per vertex for each germplasm and each location.
```{r,warning=FALSE}
p_bar = plot(net_bipart, plot_type = "barplot")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2005-2006-2007-2008-2009`$barplot
p_bar$germplasm
p_bar$location
```


Location present on the network can be displayed on a map with `plot_type = "map"`.

```{r,warning=FALSE,message=FALSE}
p_map = PPBstats:::plot.data_network(net_bipart[1], plot_type = "map", labels_on = "location")
p_map$`2005-2006-2007-2008-2009`
```

As well as plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1$`2005-2006-2007-2008-2009`
```

```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2$`2005-2006-2007-2008-2009`
```

