## Bipart network analysis
This section deals with bipart network that represents "which location has which germplasm which year": a data frame with the followin compulsory columns: `"germplasm"`, `"location"`, `"year"`.
Possible options are : `"long"`, `"lat"` to get map representation

The format of the data are checked by the function `format_data_PPBstats()`. 

The following arguments can be used :

- `type` : `"data_network"`
- `network_part` : `"bipart"`
- `vertex_type` : `c("germplasm", "location")`

Possible values of arguments regarding network formats are displayed in Table \@ref(tab:network-format-arg).

| | `network_part` | `vertex_type` | `network_split` | 
| --- | --- | --- | --- | 
| unipart for seed lots format |  `unipart` or  `bipart` |  `seed_lots` or `location` or  `c("germplasm", "location")` |  `NULL` or `germplasm` or `relation_year_start` | 
| unipart for location format |  `unipart` |  `location` |  `germplasm` or `relation_year_start` |
| bipart for germplasm and location format |  `bipart` |  `c("germplasm", "location")` |  `NULL` | 
Table: (\#tab:network-format-arg) Possible values of argument (in colum) regarding network format (in row).

The following sections give examples for each network format.
The function returns list of `igraph` object^[http://igraph.org/r/] coming from `igraph::graph_from_data_frame()`.

### Format data
```{r}
data(data_network_unipart_sl)
head(data_network_unipart_sl)
```

```{r}
net_bipart = format_data_PPBstats(
  type = "data_network",
  data = data_network_unipart_sl, 
  network_part = "bipart", 
  vertex_type =  c("germplasm", "location")
  )
```

For bipart network, it returns a list with as many elements as year in the data 
as well as all years merged in the first element of the list.
If no year are provided into the data, all information are merged.

```{r}
names(net_bipart)
```

### Describe the data

The different representations are done with the `plot` function.

```{r,warning=FALSE}
p_net = plot(net_bipart, plot_type = "network", 
                          labels_on = TRUE, labels_size = 4)
names(p_net) # one element per year, the first element with all the data
p_net$`2009`
```


With barplots, it represents the number of edges per vertex for each germplasm and each location.
```{r,warning=FALSE}
p_bar = plot(net_bipart, plot_type = "barplot")
names(p_bar) # one element per year, the first element with all the data
p_bar = p_bar$`2005-2006-2007-2008-2009`$barplot
p_bar$germplasm
p_bar$location
```


Location present on the network can be displayed on a map with `plot_type = "map"`.

```{r,warning=FALSE,message=FALSE}
p_map = PPBstats:::plot.data_network(net_bipart[1], plot_type = "map", labels_on = "location")
p_map$`2005-2006-2007-2008-2009`
```

As well as plot information regarding a variable on map with a pie with `plot_type = "map"` and by setting arguments `data_to_pie` and `vec_variables`:

```{r,warning=FALSE,message=FALSE}
# y1 is a quantitative variable
p_map_pies_y1 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y1")
p_map_pies_y1$`2005-2006-2007-2008-2009`
```

```{r,warning=FALSE,message=FALSE}
# y2 is a qualitative variable
p_map_pies_y2 = PPBstats:::plot.data_network(net_bipart[1], data_to_pie, plot_type = "map", vec_variables = "y2")
p_map_pies_y2$`2005-2006-2007-2008-2009`
```

Note that it is not possible to display plot with `plot_type = "network"`.


