## Describe the data {#describe-data-agro}

Once the data have been collected, a first step is to describe them with `describe_data`.
Seven types of plot, throught the `plot_type` argument are possible:

- presence abscence matrix that represent the combinaison of germplasm $\times$ location
- histogramm
- barplot, where sd error are displayed
- boxplot
- interaction
- biplot
- radar

Then you must choose which factor to represent on the x axis (`x_axis` argument),
the factor to display in color (`in_col` argument), and of course the variables to describe (`vec_variables` argument).

It is possible to tune the number of factor displayed (`nb_parameters_per_plot_x_axis` and `nb_parameters_per_plot_in_col` arguments) and the size of the labels regarding biplot and radar (`labels_on` and `labels_size` arguments).

### Format the data

Get two data set to look at some examples
```{r}
data("data_GxE")
data_GxE = format_data_PPBstats(data_GxE, type = "data_agro")

data("data_model_2")
data_model_2 = format_data_PPBstats(data_model_2, type = "data_agro")
```


### presence abscence matrix

The presence absence matrix may be different from experimental design planned because of NA.
The plot represents the presence/absence matrix of G $\times$ E combinations.

```{r}
p = plot(
  data_GxE, plot_type = "pam",
  vec_variables = c("y1", "y2")
  )
names(p)
p$y1
```

A score of 3 is for a given germplasm replicated three times in a given environement.

```{r}
p = plot(
  data_model_2, plot_type = "pam",
  vec_variables = c("y1", "y2")
  )
p$y1
```

Here there are lots of 0 meaning that a lot of germplasm are no in at least two locations.
A score of 1 is for a given germplasm in a given location.
A score of 2 is for a given germplasm replicated twice in a given location.


### histogramm

```{r}
p = plot(
  data_GxE, plot_type = "histogramm",
  vec_variables = c("y1", "y2")
  )
p$y1
```


### barplot

```{r}
p = plot(
  data_GxE, plot_type = "barplot",
  vec_variables = c("y1", "y2"),
  x_axis = "germplasm"
  )
```

Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_x_axis` parameters per graph.

```{r}
names(p$y1)
p$y1$`germplasm-1|-NA`
```

```{r}
p = plot(
  data_GxE, plot_type = "barplot",
  vec_variables = c("y1", "y2"),
  x_axis = "germplasm",
  in_col = "location"
  )
```

Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_x_axis` and `nb_parameters_per_in_col` parameters per graph.
```{r}
names(p$y1)
p$y1$`germplasm-1|location-1`
```


### boxplot

```{r}
p = plot(
  data_GxE, plot_type = "boxplot",
  vec_variables = c("y1", "y2"),
  x_axis = "germplasm"
  )
```

Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_x_axis` parameters per graph.

```{r}
names(p$y1)
p$y1$`germplasm-1|-NA`
```

```{r}
p = plot(
  data_GxE, plot_type = "boxplot",
  vec_variables = c("y1", "y2"),
  x_axis = "germplasm",
  in_col = "location"
  )
```

Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_x_axis` and `nb_parameters_per_in_col` parameters per graph.
```{r}
names(p$y1)
p$y1$`germplasm-1|location-1`
```


### interaction

```{r}
p = plot(
  data_GxE, plot_type = "interaction",
  vec_variables = c("y1", "y2"),
  x_axis = "germplasm",
  in_col = "location"
  )
```

Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_x_axis` and `nb_parameters_per_in_col` parameters per graph.
```{r}
names(p$y1)
p$y1$`germplasm-1|location-1`
```

It is also possible to have on the `x_axis` the date in julian day that have been automatically calculated from `format_data_PPBstats()`. 
Note that this is possible only for `plot_type = "histogramm"`, `"barplot"`, `"boxplot"` and `"interaction"`.

```{r}
p = plot.data_agro(
  data_GxE, plot_type = "interaction",
  vec_variables = c("y1", "y2"),
  x_axis = "date_julian",
  in_col = "location"
)
p$y1$`y1$date_julian-1|location-1`
```



### biplot

```{r}
p = plot(
  data_GxE, plot_type = "biplot",
  vec_variables = c("y1", "y2", "y3"),
  in_col = "germplasm", labels_on = "germplasm"
  )
```

The name of the list correspond to the pairs of variables displayed.
Note that for each element of the following list, there are as many graph as needed with `nb_parameters_per_in_col` parameters per graph.
```{r}
names(p)
p$`y1 - y2`$`-NA|germplasm-1`
```


### radar

```{r}
p = plot(
  data_GxE, plot_type = "radar",
  vec_variables = c("y1", "y2", "y3"),
  in_col = "location"
  )
p
```
