### Classic ANOVA (**M4a**) {#classic-anova}


#### Theory of the model

The experimental design used is fully replicated (**D1**) on one location.
The model is based on frequentist statistics (section \@ref(section-freq)).
The tests to check the model are explained in section \@ref(check-model-freq).
The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).
The analysis is done the following model :

$Y_{ijk} = \mu + \alpha_{i} + rep_{k} + \varepsilon_{ijk}; \quad \varepsilon_{ijk} \sim \mathcal{N} (0,\sigma^2)$

With,

- $Y_{ijk}$ the phenotypic value for replication $k$, germplasm $i$ and individual $j$,
- $\mu$ the general mean,
- $\alpha_{i}$ the effect of germplasm $i$,
- $rep_{k}$ the effect of the replication $k$,
- $\varepsilon_{ijk}$ the residuals.


#### Steps with `PPBstats`

For classic anova analysis, you can follow these steps (Figure \@ref(fig:main-workflow)):

- Format the data with `format_data_PPBstats()`
- Run the model with `model_anova()`
- Check model outputs to know if you can continue the analysis with `check_model()`  and vizualise it with `plot()`
- Get mean comparisons for each factor with `mean_comparisons()` and vizualise it with `plot()`

#### Format the data

A subset of `data_model_GxE` is used in this exemple.

```{r}
data(data_model_GxE)
data_model_anova = droplevels(dplyr::filter(data_model_GxE, location == "loc-1"))  
data_model_anova = format_data_PPBstats(data_model_anova, type = "data_agro")
head(data_model_anova)
```


#### Run the model
To run model on the dataset, used the function `model_anova`.
You can run it on one variable.

```{r}
out_anova = model_anova(data_model_anova, variable = "y1")
```

`out_anova` is a list containing two elements :

- `info` : a list with variable
```{r}
out_anova$info
```
- `ANOVA` a list with two elements :
    - `model`
    ```{r}
    out_anova$ANOVA$model
    ```
    - `anova_model`
    ```{r}
    out_anova$ANOVA$anova_model
    ```
    - `germplasm_effects` a list of two elements :
        - `effects`
        ```{r}
        out_anova$ANOVA$germplasm_effects$effects
        ```
        - `intra_variance`
        ```{r}
        out_anova$ANOVA$germplasm_effects$intra_variance
        ```

#### Check and visualize model outputs

The tests to check the model are explained in section \@ref(check-model-freq).

##### Check the model

Once the model is run, it is necessary to check if the outputs can be taken with confidence.
This step is needed before going ahead in the analysis (in fact, object used in the next functions must come from `check_model()`).

```{r}
out_check_anova = check_model(out_anova)
```

`out_check_anova` is a list containing four elements :

- `info` : a list with variable
- `model_anova` the output from the model
- `data_ggplot` a list containing information for ggplot:
    - `data_ggplot_residuals` a list containing :
        - `data_ggplot_normality`
        - `data_ggplot_skewness_test`
        - `data_ggplot_kurtosis_test`
        - `data_ggplot_shapiro_test`
        - `data_ggplot_qqplot`
    - `data_ggplot_variability_repartition_pie`
    - `data_ggplot_var_intra`

##### Visualize outputs

Once the computation is done, you can visualize the results with `plot()`

```{r}
p_out_check_anova = plot(out_check_anova)
```

`p_out_check_anova` is a list with:

- `residuals`
    - `histogram` : histogram with the distribution of the residuals
    ```{r}
    p_out_check_anova$residuals$histogram
    ```
    - `qqplot`
    ```{r}
    p_out_check_anova$residuals$qqplot
    ```
    - `points`
    ```{r}
    p_out_check_anova$residuals$points
    ```

- `variability_repartition` : pie with repartition of SumSq for each factor
```{r}
p_out_check_anova$variability_repartition
```

- `variance_intra_germplasm` : repartition of the residuals for each germplasm (see Details for more information)
With the hypothesis than the micro-environmental variation is equaly distributed on all the individuals (i.e. all the plants), the distribution of each germplasm represent the intra-germplasm variance.
This has to been seen with caution:
    - If germplasm have no intra-germplasm variance (i.e. pure line or hybrides) then the distribution of each germplasm represent only the micro-environmental variation.
    - If germplasm have intra-germplasm variance (i.e. population such as landraces for example) then the distribution of each germplasm represent the micro-environmental variation plus the intra-germplasm variance.


```{r}
p_out_check_anova$variance_intra_germplasm
```


#### Get and visualize mean comparisons

The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).

##### Get mean comparisons

Get mean comparisons with `mean_comparisons()`.

```{r}
out_mean_comparisons_anova = mean_comparisons(out_check_anova, p.adj = "bonferroni")
```

`out_mean_comparisons_anova` is a list of two elements:

- `info` : a list with variable
- `data_ggplot_LSDbarplot_germplasm`

##### Visualize mean comparisons

```{r}
p_out_mean_comparisons_anova = plot(out_mean_comparisons_anova)
```

`p_out_mean_comparisons_anova` is a list of one element with barplots :

For each element of the list, there are as many graph as needed with `nb_parameters_per_plot` parameters per graph.
Letters are displayed on each bar. Parameters that do not share the same letters are different regarding type I error (alpha) and alpha correction. 
The error I (alpha) and the alpha correction are displayed in the title. 

- `germplasm` : mean comparison for germplasm
```{r}
pg = p_out_mean_comparisons_anova$germplasm
names(pg)
pg$`1`
```

#### Get and vizualise groups of parameters

##### Get groups of parameters

In order to cluster locations or germplasms, you may use mulivariate analysis on a matrix with several variables in columns and parameter in rows.

This is done with `parameter_groups()` which do a PCA on this matrix.

Clusters are done based on HCPC method as explained [here](http://www.sthda.com/english/wiki/hcpc-hierarchical-clustering-on-principal-components-hybrid-approach-2-2-unsupervised-machine-learning) 


Lets' have an example with three variables.

First run the models
```{r}
out_anova_2 = model_anova(data_model_anova, variable = "y2")
out_anova_3 = model_anova(data_model_anova, variable = "y3")
```

Then check the models
```{r}
out_check_anova_2 = check_model(out_anova_2)
out_check_anova_3 = check_model(out_anova_3)
```

Then run the function for germplasm.

```{r}
out_parameter_groups = parameter_groups(
  list("y1" = out_check_anova, "y2" = out_check_anova_2, "y3" = out_check_anova_3), 
  "germplasm"
  )
```

`out_parameter_groups` is list of two elements:

- `obj.pca` : the PCA object from `FactoMineR::PCA()`
- `clust`, a list of two elements:
    - `res.hcpc` : the HCPC object from `FactoMineR::HCPC()`
    - `clust` : the dataframe with cluster assigned to each individual

##### Visualize groups of parameters

Visualize outputs with `plot`

```{r}
p_germplasm_group = plot(out_parameter_groups)
```

`p_germplasm_group` is list of two elements :

- `pca` : a list with three elements on the PCA on the group of parameters :

    - `composante_variance` : variance caught by each dimension of the PCA
    ```{r}
    p_germplasm_group$pca$composante_variance
    ```

    - `ind` : graph of individuals
    ```{r}
    p_germplasm_group$pca$ind
    ```

    - `var` : graph of variables
    ```{r}
    p_germplasm_group$pca$var
    ```
  
- `clust` : output from `factextra::fviz_cluster()`, a list of number of cluster + 1 element
```{r}  
cl = p_germplasm_group$clust
names(cl)
```

```{r}  
cl$cluster_all
```

```{r}
cl$cluster_1
```


#### post hoc analysis to visualize variation repartition for several variables


```{r}
list_out_check_model = list("anova_1" = out_check_anova, "anova_2" = out_check_anova_2, "anova_3" = out_check_anova_3)
post_hoc_variation(list_out_check_model)

```



#### Apply the workflow to several variables

If you wish to apply the AMMI workflow to several variables, you can use `lapply()` with the following code :

```{r}
workflow_anova = function(x, data){
  out_anova = model_anova(data, variable = x)
  
  out_check_anova = check_model(out_anova)
  p_out_check_anova = plot(out_check_anova)
  
  out_mean_comparisons_anova = mean_comparisons(out_check_anova, p.adj = "bonferroni")
  p_out_mean_comparisons_anova = plot(out_mean_comparisons_anova)
  
  out = list(
    "out_anova" = out_anova,
    "out_check_anova" = out_check_anova,
    "p_out_check_anova" = p_out_check_anova,
    "out_mean_comparisons_anova" = out_mean_comparisons_anova,
    "p_out_mean_comparisons_anova" = p_out_mean_comparisons_anova
  )
  
  return(out)
}

vec_variables = c("y1", "y2", "y3")

out = lapply(vec_variables, workflow_anova, data_model_anova)
names(out) = vec_variables

list_out_check_model = list("anova_1" = out$y1$out_check_anova, "anova_2" = out$y2$out_check_anova, "anova_3" = out$y3$out_check_anova)

out_parameter_groups = parameter_groups(list_out_check_model, "germplasm" )
p_germplasm_group = plot(out_parameter_groups)

p_post_hoc_variation = post_hoc_variation(list_out_check_model)

```


