### Spatial analysis (**M4b**) {#spatial-analysis}

Note that this section is still under development : cf https://github.com/priviere/PPBstats/issues/20

#### Theory of the model

The experimental design used is the row-column design (**D3**).
The following model is based on frequentist statistics (section \@ref(section-freq)).
The model allows taking into account environmental variation within a block with few control replicated in rows and columns.

It is based on a SpATS (Spatial Analysis of Field Trials with Splines) model proposed by @rodriguez_spatial_2016 :

$Y_{ijk} = \alpha_{i} + r_{j} + c_{k} + f(u,v) + \varepsilon_{ijk}; \quad \varepsilon_{ijk} \sim \mathcal{N} (0,\sigma^2)$

With,

- $Y_{ijk}$ the phenotypic value for germplasm $i$, row $j$ and column $k$
- $\alpha_{i}$ the random effect of germplasm $i$
- $r_{j}$ the effect of row $j$ 
- $c_{k}$ the effect of col $k$ 
- $f(u,v)$ the smooth bivariate function that simultaneously accounts for the spatial trend across both directions of the fiel (i.e. rows and columns)
- $\varepsilon_{ijk}$ the residuals


Note that $f(u,v)$ is divided into 8 components excluding the intercept [@rodriguez_spatial_2016]:

- the linear effect of the rows (row),
- the linear effect of the columns (col), 
- the linear interaction of rows and columns (row:col),
- the main row effect (f(row)),
- the main column effect (f(col)),
- the smooth varying coefficient term regarding rows (f(col):row),
- the smooth varying coefficient term regarding columns (row:f(col))),
- the smooth-by-smooth interaction component (f(col):f(row))

Much more information regarding the model as well as example of R package `SpATS` can be found in @rodriguez_spatial_2016.


#### Steps with `PPBstats`

- Format the data with `format_data_PPBstats()`
- Run the model with `model_spatial()`
- Check model outputs with graphs to know if you can continue the analysis with `check_model()` and vizualise it with `plot()`
- Get mean comparisons for germplasms with `mean_comparisons()` and vizualise it with `plot()`

#### Format the data

```{r}
p_case2 = design_experiment(
  location = "Location-3",
  year = "2016",
  expe.type = "row-column",
  germplasm = paste("germ", c(1:20), sep = ":"),
  controls = "germ:ctrl",
  nb.controls.per.block = 7,
  nb.blocks = 1,
  nb.cols = 7)

d = p_case2$`row-column`$data.frame
d$variable = as.numeric(rnorm(nrow(d), 50, 10))
d = format_data_PPBstats(d, type = "data_agro")
```


#### Run the model

By default, germplasm are settled as random (`genotype.as.random = TRUE`).

```{r}
out_spatial = model_spatial(data = d, variable = "variable")
```

`out_spatial` is a list containing two elements :

- `info` : a list with variable and data
```{r}
out_spatial$info$variable

head(out_spatial$info$data)
```
- `model` a list with five elements :
    - `model`
    ```{r}
    out_spatial$model$model
    ```
    - `summary`
    ```{r}
    out_spatial$model$summary
    ```
    - `germplasm_effects`
    ```{r}  
    out_spatial$model$germplasm_effects
    ```
    - `var_res` the variance of the residuals
    ```{r}
    out_spatial$model$df_residual
    ```
    ```{r}
    out_spatial$model$MSerror
    ```

#### Check and visualize model outputs

The tests to check the model are explained in section \@ref(check-model-freq).

##### Check the model

```{r}
out_check_spatial = check_model(out_spatial)
```

`out_check_spatial` is a list containing two elements :

- `spatial` the output from the model
- `data_ggplot` a list containing information for ggplot:
    - `data_ggplot_residuals` a list containing :
        - `data_ggplot_normality`
        - `data_ggplot_skewness_test`
        - `data_ggplot_kurtosis_test`
        - `data_ggplot_qqplot`
    - `data_ggplot_variability_repartition_pie`


##### Visualize outputs

Once the computation is done, you can visualize the results with `plot`

```{r}
p_out_check_spatial = plot(out_check_spatial)
```

`p_out_check_spatial` is a list with:

- `residuals`
    - `histogram` : histogram with the distribution of the residuals
    ```{r}
    p_out_check_spatial$residuals$histogram
    ```
    - `qqplot`
    ```{r}
    p_out_check_spatial$residuals$qqplot
    ```

- `variability_repartition` : pie with repartition of variance for each factor
```{r}
p_out_check_spatial$variability_repartition
```

#### Get and visualize mean comparisons

The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).

#### Get mean comparisons

Get mean comparisons with `mean_comparisons`.

```{r}
out_mean_comparisons_spatial = mean_comparisons(out_check_spatial, p.adj = "bonferroni")
```

`out_mean_comparisons_spatial` is a list of two elements:

- `info` : a list with variable and data
- `data_ggplot_LSDbarplot_germplasm`

#### Visualize mean comparisons
```{r}
p_out_mean_comparisons_spatial = plot(out_mean_comparisons_spatial)
```

`p_out_mean_comparisons_spatial` is a list of one elements with barplots :

For each element of the list, there are as many graph as needed with `nb_parameters_per_plot` parameters per graph.
Letters are displayed on each bar. Parameters that do not share the same letters are different regarding type I error (alpha) and alpha correction. 
The error I (alpha) and the alpha correction are displayed in the title. 

`germplasm` : mean comparison for germplasm
```{r}
pg = p_out_mean_comparisons_spatial$germplasm
names(pg)
pg$`1`
```

#### Apply the workflow to several variables

If you wish to apply the spatial workflow to several variables, you can use `lapply` with the following code :

```{r}
workflow_spatial = function(x){
  out_spatial = model_spatial(data = d, variable = "variable")
  
  out_check_spatial = check_model(out_spatial)
  p_out_check_spatial = plot(out_check_spatial)
  
  out_mean_comparisons_spatial = mean_comparisons(out_check_spatial, p.adj = "bonferroni")
  p_out_mean_comparisons_spatial = plot(out_mean_comparisons_spatial)
  
  out = list(
    out_spatial = out_spatial,
    out_check_spatial = out_check_spatial,
    p_out_check_spatial = p_out_check_spatial,
    out_mean_comparisons_spatial = out_mean_comparisons_spatial,
    p_out_mean_comparisons_spatial = p_out_mean_comparisons_spatial
    )
  return(out)
}

## Not run 
# vec_variables = c("y1", "y2", "y3")
#
# out = lapply(vec_variables, workflow_spatial)
# names(out) = vec_variables
```


