### AMMI {#ammi}

#### Theory of the model

The model is based on frequentist statistics (section \@ref(section-freq)).
The Additive Main effects and Multiplicative Interaction (AMMI) model is based on two analyses [@gauch_statistical_2006] : 


1. an **ANOVA** with the following model :

$Y_{ijk} = \mu + \alpha_{i} + \theta_{j} + rep_{k}(\theta_{j}) + \eta_{i}\nu_{j} + \varepsilon_{ijk}; \quad \varepsilon_{ijk} \sim \mathcal{N} (0,\sigma^2)$

With,

- $Y_{ijk}$ the phenotypic value for replication $k$, germplasm $i$ and location $j$,
- $\mu$ the general mean,
- $\alpha_{i}$ the germplasm $i$ effect,
- $\theta_{j}$ location $j$ effect,
- $rep_{k}(\theta_{j})$ the replication $k$ nested in location effect,
- $\eta_{i}\nu_{j}$ the germplasm $\times$ location effect,
- $\varepsilon_{ijk}$ the residuals.


Or, if there are several years in the data set:

$Y_{ijkl} = \mu + \alpha_{i} + \theta_{j} + rep_{k}(\theta_{j}\beta_{l}) + \eta_{i}\nu_{j} + \beta_{l} + \beta_{l}\alpha_{i} + \beta_{l}\theta_{j} + \varepsilon_{ijk}; \quad \varepsilon_{ijk} \sim \mathcal{N} (0,\sigma^2)$

With,

- $Y_{ijkl}$ the phenotypic value for replication $k$, germplasm $i$, location $j$ and year $l$,
- $\mu$ the general mean,
- $\alpha_{i}$ the germplasm $i$ effect,
- $\theta_{j}$ location $j$ effect,
- $rep_{k}(\theta_{j}\beta_{l})$ the replication $k$ nested in location and year effect,
- $\eta_{i}\nu_{j}$ the germplasm $\times$ location effect,
- $\beta_{l}$ the year $l$ effect,
- $\beta_{l}\alpha_{i}$ the year $\times$ germplasm interaction effect,
- $\beta_{l}\theta_{j}$ the year $\times$ location interaction effect,
- $\varepsilon_{ijk}$ the residuals.


2. a **PCA** which analyse deeper the germplasm $\times$ location interaction : 

$\eta_{i}\nu_{j} = \sum_{n}^{N} \lambda_{n} \gamma_{in} \omega_{jn}$ ^[Note than in `PPBstats`, the PCA is done on $\eta_{i}\nu_{j} + \varepsilon_{ijk}$. The residuals is in the last dimension of the PCA]


which can also be written :

$\eta_{i}\nu_{j} = \sum_{n}^{N} (\sqrt{\lambda_{n}} \gamma_{in}) (\sqrt{\lambda_{n}} \omega_{jn})$

With,

- $\eta_{i}\nu_{j}$ interaction of germplasm $i$ with location $j$,
- $N$ the number of dimension (PCA componant) which has as maximum value the number of location,
- $\lambda_{n}$ the eigen value for componant $n$,
- $\gamma_{in}$ eigen vector for germplasm $i$ for componant $n$,
- $\omega_{jn}$ eigen vector for  location $j$  for componant $n$.

The data are double centered on location and germplasm.
The PCA study the structure of the interaction matrix.
The location are the variable and the germplasm are the individuals.

This PCA allows to detect

- germplasm that are stable (i.e. contribute less to interaction)
- which germplasm interact the most with which location
- which location have the same profile regarding interaction


#### Steps with `PPBstats`

For AMMI analysis, you can follow these steps (Figure \@ref(fig:main-workflow)):

- Format the data with `format_data_PPBstats()`
- Run the model with `GxE()` and `gxe_analysis = "AMMI"`
- Check model outputs with graphs to know if you can continue the analysis with `check_model()`
- Get mean comparisons for each factor with `mean_comparisons()` and vizualise it with `plot()`
- Get groups of each parameters with `parameters_groups()` and visualise it with `plot()`

#### Format the data

```{r}
data(data_GxE)
data_GxE = format_data_PPBstats(data_GxE, type = "data_agro")
head(data_GxE)
```


#### Run the model
To run model GxE on the dataset, used the function `GxE`.
You can run it on one variable.

```{r}
out_ammi = GxE(data_GxE, variable = "y1", gxe_analysis = "AMMI")
```

`out_ammi` is a list containing three elements :

- `info` : a list with variable and gxe\_analysis
```{r}
out_ammi$info
```
- `ANOVA` a list with five elements :
    - `model`
    ```{r}
    out_ammi$ANOVA$model
    ```
    - `anova_model`
    ```{r}
    out_ammi$ANOVA$anova_model
    ```
    - `germplasm_effects` a list of two elements :
        - `effects`
        ```{r}
        out_ammi$ANOVA$germplasm_effects$effects
        ```
        - `intra_variance`
        ```{r}
        out_ammi$ANOVA$germplasm_effects$intra_variance
        ```

    - `location_effects`
    ```{r}
    out_ammi$ANOVA$location_effects$effects
    ```
    - `interaction_matrix`
    ```{r}
    out_ammi$ANOVA$interaction_matrix
    ```
- `PCA` : PCA object from FactoMineR
```{r}
out_ammi$PCA
```

#### Check and visualize model outputs

The tests to check the model are explained in section \@ref(check-model-freq).

##### Check the model

Once the model is run, it is necessary to check if the outputs can be taken with confidence.
This step is needed before going ahead in the analysis (in fact, object used in the next functions must come from `check_model()`).

```{r}
out_check_ammi = check_model(out_ammi)
```

`out_check_ammi` is a list containing four elements :

- `GxE` the output from the model
- `data_ggplot` a list containing information for ggplot:
    - `data_ggplot_residuals` a list containing :
        - `data_ggplot_normality`
        - `data_ggplot_skewness_test`
        - `data_ggplot_kurtosis_test`
        - `data_ggplot_qqplot`
    - `data_ggplot_variability_repartition_pie`
    - `data_ggplot_var_intra`

##### Visualize outputs

Once the computation is done, you can visualize the results with `plot`

```{r}
p_out_check_ammi = plot(out_check_ammi)
```

`p_out_check_ammi` is a list with:

- `residuals`
    - `histogram` : histogram with the distribution of the residuals
    ```{r}
    p_out_check_ammi$residuals$histogram
    ```
    - `qqplot`
    ```{r}
    p_out_check_ammi$residuals$qqplot
    ```

- `variability_repartition` : pie with repartition of SumSq for each factor
```{r}
p_out_check_ammi$variability_repartition
```

- `variance_intra_germplasm` : repartition of the residuals for each germplasm (see Details for more information)
With the hypothesis than the micro-environmental variation is equaly distributed on all the individuals (i.e. all the plants), the distribution of each germplasm represent the intra-germplasm variance.
This has to been seen with caution:
    - If germplasm have no intra-germplasm variance (i.e. pure line or hybrides) then the distribution of each germplasm represent only the micro-environmental variation.
    - If germplasm have intra-germplasm variance (i.e. population such as landraces for example) then the distribution of each germplasm represent the micro-environmental variation plus the intra-germplasm variance.


```{r}
p_out_check_ammi$variance_intra_germplasm
```

- `pca_composante_variance` : variance caught by each dimension of the PCA run on the interaction matrix
```{r}
p_out_check_ammi$pca_composante_variance
```


#### Get and visualize mean comparisons

The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).

##### Get mean comparisons

Get mean comparisons with `mean_comparisons`.

```{r}
#out_mean_comparisons_ammi = mean_comparisons(out_check_ammi, p.adj = "bonferroni")
```

`out_mean_comparisons_ammi` is a list of three elements:

- `info` : a list with variable and gxe_analysis
- `data_ggplot_LSDbarplot_germplasm`
- `data_ggplot_LSDbarplot_location`
- `data_ggplot_LSDbarplot_year`

##### Visualize mean comparisons

```{r}
# p_out_mean_comparisons_ammi = plot(out_mean_comparisons_ammi)
```

`p_out_mean_comparisons_ammi` is a list of three elements with barplots :

For each element of the list, there are as many graph as needed with `nb_parameters_per_plot` parameters per graph.
Letters are displayed on each bar. Parameters that do not share the same letters are different regarding type I error (alpha) and alpha correction. 
The error I (alpha) and the alpha correction are displayed in the title. 

- `germplasm` : mean comparison for germplasm
```{r}
#pg = p_out_mean_comparisons_ammi$germplasm
#names(pg)
#pg$`1`
```

- `location` : mean comparison for location
```{r}
#pl = p_out_mean_comparisons_ammi$location
#names(pl)
#pl$`1`
```

- `year` : mean comparison for year
Here there are no more than 2 years in the data set so it is `NULL`
```{r}
#p_out_mean_comparisons_ammi$year
```

#### Get and visualize biplot

The biplot represents information about the percentages of total variation explained by the two axes.
It has to be linked to the total variation caught by the interaction.
If the total variation is small, then the biplot is useless.
If the total variation is high enought, then the biplot is useful if the two first dimension represented catch enought variation (the more the better).

##### Get biplot

```{r}
out_biplot_ammi = biplot_GxE(out_check_ammi)
```

##### Visualize biplot

```{r}
p_out_biplot_ammi = plot(out_biplot_ammi)
```

`p_out_biplot_ammi` is a list of three elements :
- `ecovalence`

Ecovalence from @wricke_uber_1962 give part of interaction variance taken by germplasm and location.
It is an indicator of stability: a low ecovalence means low interaction, i.e. more stability.

Ecovalance of germplasm $i$ is $W_{i}=\sum_{i}^{n} (\eta_{i}\nu_{j})^{2}$

Ecovalance of location $j$ is $W_{j}=\sum_{j}^{n} (\eta_{i}\nu_{j})^{2}$.

Ecovalances are represented in fonction of mean effects by germplasm and location.

```{r}
p_out_biplot_ammi$ecovalence
```

- `interaction` which displays the interaction matrix

```{r}
p_out_biplot_ammi$interaction
```


- `biplot` being a list of four elements :
```{r}
p_out_biplot_ammi$biplot$simple_biplot
```

Regarding the other elements of the list, it returns `NULL` as these visualisation is only done for `gxe_analysis = "GGE"`.

```{r}
p_out_biplot_ammi$biplot$which_won_where
p_out_biplot_ammi$biplot$mean_vs_stability
p_out_biplot_ammi$biplot$discrimitiveness_vs_representativeness
```


#### Get and vizualise groups of parameters

##### Get groups of parameters

In order to cluster locations or germplasms, you may use mulivariate analysis on a matrix with several variables in columns and parameter in rows.

This is done with `parameter_groups` which do a PCA on this matrix.

Clusters are done based on HCPC method as explained here http://www.sthda.com/english/wiki/hcpc-hierarchical-clustering-on-principal-components-hybrid-approach-2-2-unsupervised-machine-learning

Lets' have an example with three variables.

First run the models
```{r}
out_ammi_2 = GxE(data_GxE, variable = "y2", gxe_analysis = "AMMI")
out_ammi_3 = GxE(data_GxE, variable = "y3", gxe_analysis = "AMMI")
```

Then check the models
```{r}
out_check_ammi_2 = check_model(out_ammi_2)
out_check_ammi_3 = check_model(out_ammi_3)
```

Then run the function for germplasm. It can also be done on location or year if there are year in the data set.

```{r}
out_parameter_groups = parameter_groups(
  list("y1" = out_check_ammi, "y2" = out_check_ammi_2, "y3" = out_check_ammi_3), 
  "germplasm"
  )
```

`out_parameter_groups` is list of two elements:

- `obj.pca` : the PCA object from `FactoMineR::PCA`
- `clust`, a list of two elements:
    - `res.hcpc` : the HCPC object from `FactoMineR::HCPC`
    - `clust` : the dataframe with cluster assigned to each individual

##### Visualize groups of parameters

Visualize outputs with `plot`

```{r}
p_germplasm_group = plot(out_parameter_groups)
```

`p_germplasm_group` is list of two elements :

- `pca` : a list with three elements on the PCA on the group of parameters :

    - `composante_variance` : variance caught by each dimension of the PCA
    ```{r}
    p_germplasm_group$pca$composante_variance
    ```

    - `ind` : graph of individuals
    ```{r}
    p_germplasm_group$pca$ind
    ```

    - `var` : graph of variables
    ```{r}
    p_germplasm_group$pca$var
    ```
  
- `clust` : output from `factextra::fviz_cluster()`, a list of number of cluster + 1 element
```{r}  
cl = p_germplasm_group$clust
names(cl)
```

```{r}  
cl$cluster_all
```

```{r}
cl$cluster_1
```

#### Apply the workflow to several variables {#workflow_gxe}

If you wish to apply the AMMI workflow to several variables, you can use `lapply` with the following code :

```{r}
workflow_gxe = function(x, gxe){
  out_gxe = GxE(data_GxE, variable = x, gxe_analysis = gxe)
  
  out_check_gxe = check_model(out_gxe)
  p_out_check_gxe = plot(out_check_gxe)
  
  out_mean_comparisons_gxe = mean_comparisons(out_check_gxe, p.adj = "bonferroni")
  p_out_mean_comparisons_gxe = plot(out_mean_comparisons_gxe)
  
  out_biplot_gxe = biplot_GxE(out_check_gxe)
  p_out_biplot_gxe = plot(out_biplot_gxe)

  out = list(
    "out_gxe" = out_gxe,
    "out_check_gxe" = out_check_gxe,
    "p_out_check_gxe" = p_out_check_gxe,
    "out_mean_comparisons_gxe" = out_mean_comparisons_gxe,
    "p_out_mean_comparisons_gxe" = p_out_mean_comparisons_gxe,
    "out_biplot_gxe" = out_biplot_gxe,
    "p_out_biplot_gxe" = p_out_biplot_gxe
  )
  
  return(out)
}

vec_variables = c("y1", "y2", "y3")

out = lapply(vec_variables, workflow_gxe, "AMMI")
names(out) = vec_variables

out_parameter_groups = parameter_groups(
  list("y1" = out$y1$out_check_gxe, "y2" = out$y2$out_check_gxe, "y3" = out$y3$out_check_gxe), 
  "germplasm" )

p_germplasm_group = plot(out_parameter_groups)

```


