### Study local adaptation

#### Home away {#home-away}

Home away analysis allows to study local adaptation.
Away in a location refers to a germplasm that has not been grown or selected in a given location.
Home in a location refers to a germplasm that has been grown or selected in a given location.

The following model take into account germplasm and location effects in order to better study version (home or away) effect [@blanquart_practical_2013].
The model is based on frequentist statistics (section \@ref(section-freq)).

$Y_{ijkm} = \mu + \alpha_i + \theta_j + \omega_{k_{ij}} + (\omega \times \alpha)_{k_{ij}j} + rep(\theta)_{mj} + \varepsilon_{ijkm}; \quad \varepsilon_{ijkm} \sim \mathcal{N} (0,\sigma^2)$

with

- $Y_{ijkm}$ the phenotypic value for replication $m$, germplasm $i$ and location $j$, and version $k$,
- $\mu$ the general mean,
- $\alpha_i$ the effect of germplasm $i$,
- $\theta_j$ the effect of location $j$,
- $\omega_{k_{ij}}$ the effect of version, home or away for a germplasm $i$ in location $j$,
- $(\omega \times \theta)_{k_{ij}j}$ the interaction effect of version $\times$ location,
- $rep(\theta)_{mj}$ the effect of the replication $m$ nested in location,
- $\varepsilon_{ijkm}$ the residuals.

The comparisons of all germplasm in all location in sympatric or allopatric situation (measured by version effect $\omega_{k_{ij}}$) give a glocal measure of local adaptation [@blanquart_practical_2013].
Interaction effect $(\omega \times \alpha)_{k_{ij}j}$ give information on specific adaptation to each location.

If there are more than one year, then the model can be written :

$Y_{ijklm} = \mu + \alpha_i + \theta_j + \beta_l + \omega_{k_{ij}} + (\theta \times \beta)_{jl} + (\omega \times \alpha)_{k_{ij}j} + rep(\theta \times \beta)_{mjl} + (\omega \times \alpha \times \beta)_{k_{ij}jl} + \varepsilon_{ijklm}; \quad \varepsilon_{ijklm} \sim \mathcal{N} (0,\sigma^2)$

with

- $Y_{ijklm}$ the phenotypic value for replication $m$, germplasm $i$ and location $j$, year $l$ and version $k$,
- $\mu$ the general mean,
- $\alpha_i$ the effect of germplasm $i$,
- $\theta_j$ the effect of location $j$,
- $\beta_l$ the effect of year $l$,
- $\omega_{k_{ij}}$ the effect of version, home or away for a germplasm $i$ in location $j$,
- $(\theta \times \beta)_{jl}$ the interaction effect of location $\times$ year,
- $(\omega \times \theta)_{k_{ij}j}$ the interaction effect of version $\times$ location,
- $rep(\theta \times \beta)_{mjl}$ the effect of the replication $m$ nested in location $\times$ year,
- $(\omega \times \theta \times \beta)_{k_{ij}jl}$ the interaction effect of version $\times$ location $\times$ year,
- $\varepsilon_{ijklm}$ the residuals.

Interaction $(\omega \times \theta \times \beta)_{k_{ij}jl}$ give information on specific adaptation to each location for a given year.

##### Steps with `PPBstats`

For home away analysis, you can follow these steps (Figure \@ref(fig:main-workflow)):

- Format the data with `format_data_PPBstats()`
- describe the data with `plot()`
- Run the model with `model_home_away()`
- Check model outputs with graphs to know if you can continue the analysis with `check_model()` and vizualise it with `plot()`
- Get mean comparisons for each factor with `mean_comparisons()` and vizualise it with `plot()`


##### Format data

```{r}
data("data_model_GxE")
data_model_GxE = format_data_PPBstats(data_model_GxE, type = "data_agro")

data("data_version_HA")
data_version_HA = format_data_PPBstats(data_version_HA, type = "data_agro_version")
head(data_version_HA)
```

Where `version` represents away or home and `group` represents the location where the germplasm come from.

##### Describe the data

```{r}
p = plot(data_model_GxE, vec_variables = "y1", data_version = data_version_HA, plot_type = "barplot")
```

`p` is a list with as many element as variable.
For each variable, there are three elements :

- A single plot with version for all germplasm merged
```{r}
p$y1$home_away_merged
```

- A single plot with version for each germplasm
```{r}
p$y1$home_away_merged_per_germplasm
```

- plot for each germplasm with all version separated
```{r}
p$y1$home_away_per_germplasm$`germ-1`
```


##### Run the model

To run HOME AWAY model on the dataset, use the function `model_home_away`.
You can run it on one variable.

```{r}
out_ha = model_home_away(data_model_GxE, data_version_HA, "y1")
```

`out_ha` is a list containing three elements :

- `info` : a list with variable
```{r}
out_ha$info
```
- `ANOVA` a list with five elements :
    - `model`
    ```{r}
    out_ha$ANOVA$model
    ```
    - `anova_model`
    ```{r}
    out_ha$ANOVA$anova_model
    ```

##### Check and visualize model outputs

The tests to check the model are explained in section \@ref(check-model-freq).


###### Check the model

Once the model is run, it is necessary to check if the outputs can be taken with confidence.
This step is needed before going ahead in the analysis (in fact, object used in the next functions must come from `check_model()`).

```{r}
out_check_ha = check_model(out_ha)
```

`out_check_ha` is a list containing four elements :

- `model_home_away` the output from the model
- `data_ggplot` a list containing information for ggplot:
    - `data_ggplot_residuals` a list containing :
        - `data_ggplot_normality`
        - `data_ggplot_skewness_test`
        - `data_ggplot_kurtosis_test`
        - `data_ggplot_qqplot`
    - `data_ggplot_variability_repartition_pie`
    - `data_ggplot_var_intra`

###### Visualize outputs

Once the computation is done, you can visualize the results with `plot()`

```{r}
p_out_check_ha = plot(out_check_ha)
```

`p_out_check_ha` is a list with:

- `residuals`
    - `histogram` : histogram with the distribution of the residuals
    ```{r}
    p_out_check_ha$residuals$histogram
    ```
    - `qqplot`
    ```{r}
    p_out_check_ha$residuals$qqplot
    ```

- `variability_repartition` : pie with repartition of SumSq for each factor
```{r}
p_out_check_ha$variability_repartition
```

- `variance_intra_germplasm` : repartition of the residuals for each germplasm (see Details for more information)
With the hypothesis than the micro-environmental variation is equaly distributed on all the individuals (i.e. all the plants), the distribution of each germplasm represent the intra-germplasm variance.
This has to been seen with caution:
    - If germplasm have no intra-germplasm variance (i.e. pure line or hybrides) then the distribution of each germplasm represent only the micro-environmental variation.
    - If germplasm have intra-germplasm variance (i.e. population such as landraces for example) then the distribution of each germplasm represent the micro-environmental variation plus the intra-germplasm variance.
```{r}
p_out_check_ha$variance_intra_germplasm
```


##### Get and visualize mean comparisons

The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).
Here, the computation is based on emmeans.

###### Get mean comparisons

Get mean comparisons with `mean_comparisons()`.

```{r}
out_mean_comparisons_ha = mean_comparisons(out_check_ha, p.adj = "tukey")
```

`out_mean_comparisons_ha` is a list of five elements:

- `info` : a list with variable
- `data_ggplot_LSDbarplot_version:germplasm`
- `data_ggplot_LSDbarplot_germplasm`
- `data_ggplot_LSDbarplot_location`
- `data_ggplot_LSDbarplot_year` in case there is year in the model

###### Visualize mean comparisons


```{r}
p_out_mean_comparisons_ha = plot(out_mean_comparisons_ha)
```

`p_out_mean_comparisons_ha` is a list of three elements with barplots :

For each element of the list, there are as many graph as needed with `nb_parameters_per_plot` parameters per graph.
Letters are displayed on each bar. Parameters that do not share the same letters are different regarding type I error (alpha) and alpha correction.
The error I (alpha) and the alpha correction are displayed in the title.

When comparing version for each germplasm, differences are displayed with stars.
The stars corresponds to the pvalue:

| pvalue | stars |
| --- | --- |
| $< 0.001$ | *** |
| $[0.001 , 0.05]$ | ** |
| $[0.05 , 0.01]$ | * |
| $> 0.01$ | . |


- `version:germplasm` : mean comparison for version for each germplasm
```{r}
pvg = p_out_mean_comparisons_ha$"version:germplasm"
pvg
```

- `germplasm` : mean comparison for germplasm
```{r}
pg = p_out_mean_comparisons_ha$germplasm
pg$`1`
```

- `location` : mean comparison for location
```{r}
pl = p_out_mean_comparisons_ha$location
pl$`1`
```

- `year` : mean comparison for year in case there is year in the model.



#### Local foreign {#local-foreign}

Another way to study local adaptation of germplasm to their location from origin is to compare germplasm behavior on their original location with their behavior on other locations : if the first is greater than the second then the germplasm is more adapted to its original location rather than to the other locations.

Local in a location refers to a germplasm that has been grown or selected in a given location.
Foreign in a location refers to a germplasm that has not been grown or selected in a given location.

The following model take into account germplasm and location effects in order to better study version (local or foreign) effect [@blanquart_practical_2013]:

$Y_{ijkm} = \mu + \alpha_i + \theta_j + \omega_{k_{ij}} + (\omega \times \theta)_{k_{ij}j} + rep(\theta)_{mj} + \varepsilon_{ijkm}; \quad \varepsilon_{ijkm} \sim \mathcal{N} (0,\sigma^2)$

with

- $Y_{ijkm}$ the phenotypic value for replication $m$, germplasm $i$ and location $j$, and version $k$,
- $\mu$ the general mean,
- $\alpha_i$ the effect of germplasm $i$,
- $\theta_j$ the effect of location $j$,
- $\omega_{k_{ij}}$ the effect of version, home or away for a germplasm $i$ in location $j$,
- (\omega \times \alpha)_{k_{ij}j} the interaction effect of version $\times$ germplasm,
- rep(\theta)_{mj} the effect of the replication $m$ nested in location,
- \varepsilon_{ijkm} the residuals.

As for home away model, version effect $\omega_{k_{ij}}$) give a glocal measure of local adaptation of germplasm to their location of origin [@blanquart_practical_2013].
Interaction effect $(\omega \times \theta)_{k_{ij}j}$ give information on specific adaptation to each germplasm.


If there are more than one year, then the model can be written :

$Y_{ijklm} = \mu + \alpha_i + \theta_j + \beta_l + \omega_{k_{ij}} + (\theta \times \beta)_{jl} + (\omega \times \alpha)_{k_{ij}i} + rep(\theta \times \beta)_{mjl} + (\omega \times \alpha \times \beta)_{k_{ij}il} + \varepsilon_{ijklm}; \quad \varepsilon_{ijklm} \sim \mathcal{N} (0,\sigma^2)$

with

- $Y_{ijklm}$ the phenotypic value for replication $m$, germplasm $i$ and location $j$, year $l$ and version $k$,
- $\mu$ the general mean,
- $\alpha_i$ the effect of germplasm $i$,
- $\theta_j$ the effect of location $j$,
- $\beta_l$ the effect of year $l$,
- $\omega_{k_{ij}}$ the effect of version, home or away for a germplasm $i$ in location $j$,
- (\theta \times \beta)_{jl} the interaction effect of location $\times$ year,
- (\omega \times \alpha)_{k_{ij}i} the interaction effect of version $\times$ germplasm,
- rep(\theta \times \beta)_{mjl} the effect of the replication $m$ nested in location $\times$ year,
- (\omega \times \alpha \times \beta)_{k_{ij}il} the interaction effect of version $\times$ germplasm $\times$ year,
- \varepsilon_{ijklm} the residuals.

Interaction $(\omega \times \theta \times \beta)_{k_{ij}jl}$ give information on specific adaptation to each germplasm for a given year.

##### Steps with `PPBstats`

For local foreign analysis, you can follow these steps (Figure \@ref(fig:main-workflow)):

- Format the data with `format_data_PPBstats()`
- describe the data with `plot()`
- Run the model with `model_local_foreign()`
- Check model outputs with graphs to know if you can continue the analysis with `check_model()` and vizualise it with `plot()`
- Get mean comparisons for each factor with `mean_comparisons()` and vizualise it with `plot()`



##### Format data
The data are formated from the home away data set

```{r}
data_version_LF = HA_to_LF(data_version_HA)
head(data_version_LF)
```

##### Describe the data

```{r}
p = plot(data_model_GxE, vec_variables = "y1", data_version = data_version_LF, plot_type = "barplot")
```

`p` is a list with as many element as variable.
For each variable, there are three elements :

- A single plot with version for all location merged
```{r}
p$y1$local_foreign_merged
```
- A single plot with version for each location
```{r}
p$y1$local_foreign_merged_per_location
```
- plot for each location with all version separated
```{r}
p$y1$local_foreign_per_location$`loc-1`
```

##### Run the model

To run LOCAL FOREIGN model on the dataset, use the function `model_local_foreign`.
You can run it on one variable.

```{r}
out_lf = model_local_foreign(data_model_GxE, data_version_LF, "y1")
```

`out_lf` is a list containing three elements :

- `info` : a list with variable
```{r}
out_lf$info
```
- `ANOVA` a list with five elements :
    - `model`
    ```{r}
    out_lf$ANOVA$model
    ```
    - `anova_model`
    ```{r}
    out_lf$ANOVA$anova_model
    ```

##### Check and visualize model outputs

The tests to check the model are explained in section \@ref(check-model-freq).


###### Check the model

Once the model is run, it is necessary to check if the outputs can be taken with confidence.
This step is needed before going ahead in the analysis (in fact, object used in the next functions must come from `check_model()`).

```{r}
out_check_lf = check_model(out_lf)
```

`out_check_lf` is a list containing four elements :

- `model_local_foreign` the output from the model
- `data_ggplot` a list containing information for ggplot:
    - `data_ggplot_residuals` a list containing :
        - `data_ggplot_normality`
        - `data_ggplot_skewness_test`
        - `data_ggplot_kurtosis_test`
        - `data_ggplot_qqplot`
    - `data_ggplot_variability_repartition_pie`
    - `data_ggplot_var_intra`

###### Visualize outputs

Once the computation is done, you can visualize the results with `plot()`

```{r}
p_out_check_lf = plot(out_check_lf)
```

`p_out_check_lf` is a list with:

- `residuals`
    - `histogram` : histogram with the distribution of the residuals
    ```{r}
    p_out_check_lf$residuals$histogram
    ```
    - `qqplot`
    ```{r}
    p_out_check_lf$residuals$qqplot
    ```

- `variability_repartition` : pie with repartition of SumSq for each factor
```{r}
p_out_check_lf$variability_repartition
```

- `variance_intra_germplasm` : repartition of the residuals for each germplasm (see Details for more information)
With the hypothesis than the micro-environmental variation is equaly distributed on all the individuals (i.e. all the plants), the distribution of each germplasm represent the intra-germplasm variance.
This has to been seen with caution:
    - If germplasm have no intra-germplasm variance (i.e. pure line or hybrides) then the distribution of each germplasm represent only the micro-environmental variation.
    - If germplasm have intra-germplasm variance (i.e. population such as landraces for example) then the distribution of each germplasm represent the micro-environmental variation plus the intra-germplasm variance.
```{r}
p_out_check_lf$variance_intra_germplasm
```


##### Get and visualize mean comparisons

The method to compute mean comparison are explained in section \@ref(mean-comp-check-freq).
Here, the computation is based on emmeans.

###### Get mean comparisons

Get mean comparisons with `mean_comparisons()`.

```{r}
out_mean_comparisons_lf = mean_comparisons(out_check_lf, p.adj = "tukey")
```

`out_mean_comparisons_lf` is a list of five elements:

- `info` : a list with variable
- `data_ggplot_LSDbarplot_version:germplasm`
- `data_ggplot_LSDbarplot_germplasm`
- `data_ggplot_LSDbarplot_location`
- `data_ggplot_LSDbarplot_year` in case there is year in the model

###### Visualize mean comparisons


```{r}
p_out_mean_comparisons_lf = plot(out_mean_comparisons_lf)
```

`p_out_mean_comparisons_lf` is a list of three elements with barplots :

For each element of the list, there are as many graph as needed with `nb_parameters_per_plot` parameters per graph.
Letters are displayed on each bar. Parameters that do not share the same letters are different regarding type I error (alpha) and alpha correction.
The error I (alpha) and the alpha correction are displayed in the title.

When comparing version for each germplasm, differences are displayed with stars.
The stars corresponds to the pvalue:

| pvalue | stars |
| --- | --- |
| $< 0.001$ | *** |
| $[0.001 , 0.05]$ | ** |
| $[0.05 , 0.01]$ | * |
| $> 0.01$ | . |


- `version:germplasm` : mean comparison for version for each germplasm
```{r}
pvg = p_out_mean_comparisons_lf$"version:germplasm"
pvg
```

- `germplasm` : mean comparison for germplasm
```{r}
pg = p_out_mean_comparisons_lf$germplasm
pg$`1`
```

- `location` : mean comparison for location
```{r}
pl = p_out_mean_comparisons_lf$location
pl$`1`
```

- `year` : mean comparison for year in case there is year in the model.
