### Steps with `PPBstats`

For variance intra analysis, you can follow these steps (Figure \@ref(fig:main-workflow)) :

- Format the data with `format_data_PPBstats()`
- Look at correlation between variables with `ggcorrplot::ggcorrplot()`^[http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2]
- Run the analysis with `multivariate()`
- Check outputs and results with functions from `factoextra`^[https://github.com/kassambara/factoextra]

### Format the data

```{r}
data("data_model_GxE")
data_model_GxE = format_data_PPBstats(data_model_GxE, type = "data_agro")
```

### Look at correlation

```{r}
library(ggcorrplot)
vec_variables = c("y1", "y2", "y3")
corr = round(cor(data_model_GxE[,vec_variables]), 1)
p.mat = cor_pmat(data_model_GxE[,vec_variables])

# Barring the no significant coefficient
ggcorrplot(corr, hc.order = TRUE, type = "lower", p.mat = p.mat)

```

More details on the use of `ggcorrplot()` can be found [here](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2).

### Run the analysis

```{r}
vec_variables = c("y1", "y2", "y3")
res.pca = mutlivariate(data_model_GxE, vec_variables, PCA)
```

### Check outputs and results

Look at the results thanks to the `factoextra` package:

```{r}
fviz_eig(res.pca)
```


```{r}
fviz_pca_ind(res.pca, label="none", habillage="location", addEllipses=TRUE, ellipse.level=0.95)
```

