# Communicate results to farmers {#com}

Within a PPB programme, it is important to give results back to the farmers in order to discuss  and accompagny them in their selection.
This can be done through the creation of a report with results coming from the analysis.

The workflow to create the report follow three steps:

1. run the analysis and create a list with all results
2. create a `.Rmd` template that will be shared by all locations
3. run `rmarkdown::render()` in order to generate report for each location


## Run the analysis

### Format data

```{r}
data("data_model_GxE")
data_model_GxE = format_data_PPBstats(data_model_GxE, type = "data_agro")
```


### Descriptive plots for each location

```{r}
vec_locations = levels(data_model_GxE$location)
list_hist_locations = lapply(
  vec_locations, function(x){
    p = plot(
      data_model_GxE, plot_type = "histogramm",
      vec_variables = c("y1")
      )
    p$y1
    }
  )
names(list_hist_locations) = vec_locations
```

Note that it is important that each element of the list refers to data of a given location in order to catch the right information in the next step when generating the report.

### GGE model for all locations

The function `workflow_gxe()` is coming from section \@ref(ammi).

```{r}
vec_variables = c("y1")
res_gge = lapply(vec_variables, workflow_gxe, "GGE")
names(res_gge) = vec_variables
```

### Create a list with all results

```{r}
res = list("hist_locations" = list_hist_locations, 
           "res_gge" = res_gge
           )
```


## Create a `.Rmd` template

Your template call two objects:

- `params` which is list with parameter of the report
- `res` which is a list with all results coming from the analysis


Below is an example of a minimal template:

```{yaml, eval=FALSE}
---
title: "`r params$title`"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Flower Seed
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---
```

````r

`r ''````{r global_options,include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

# Introduction
Here is a beautiful report with all results of the year !

# Description of data on the location

`r ''````{r}
p = res$hist_locations[[location]]
if(is.null(p)){print("No data for y1")}else {p}
```


# GGE analysis for the three locations

The germplasm effects are the following:

`r ''````{r}
res$res_gge$y1$p_out_mean_comparisons_gxe$germplasm
```

The which won where plot is the following:

`r ''````{r}
res$res_gge$y1$p_out_biplot_gxe$biplot$which_won_where
```

````

## Generate a report for each location

To generate the report, you need the `R` package `rmarkdown` installed.
In the following example, the output is `.html`. 
You can choose `.pdf` or `.docx`.
See `?rmarkdown::render` for more information.

```{r, eval=FALSE}
library(rmarkdown)

vec_locations = names(res$hist_locations)
for (location in vec_locations){ # For each location, render a report
  params = list("title" = paste("Agronomic analyses for", location))
  rmarkdown::render("./template.Rmd", 
                    output_file =  paste("report_", location, "_", ".html", sep = ""), 
                    output_dir = "."
                    )
  }

```


The report generated can be visualized for 
[loc-1](https://priviere.github.io/PPBstats_return_report_web_site/example_1_report_loc-1.html), 
[loc-2](https://priviere.github.io/PPBstats_return_report_web_site/example_1_report_loc-2.html) and 
[loc-3](https://priviere.github.io/PPBstats_return_report_web_site/example_1_report_loc-3.html).


More examples of worklows with script, template and outputs can be found [here](https://priviere.github.io/PPBstats_return_report_web_site).

