---
title: "Contribute to PPBstats"
author: "`r paste(format(eval(parse(text = devtools::as.package('.')$'authors@r')), include = c('given', 'family')), collapse = ', ')`"
date: "`r Sys.Date()`"
output:
  rmarkdown::pdf_document:
    includes: 
      in_header: header.tex
    fig_caption: yes
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\pack}{\texttt{PPBstats}}


# What to contribute

You can contribute to \pack\ in three ways:

- __test__ \pack\ and its vignette and __report__ bugs

- __translate__ \pack

- add new __functions__ that may be useful in the analysis of trials that can be
found in PPB programmes


## Test and report bugs

Do not hesitate to file [an issue in the project's website](https://github.com/priviere/PPBstats/issues)
or to send me an email if

- some points are not clear in the vignette or you find bugs

- you find bugs with your data (If so, you can send it to me so that I can see
where could be the problems)


## Translate

The package is in english. 
Every contributions to other languages are welcome!
More informations on R package translation
[here](http://developer.r-project.org/Translations30.html).

## Add new functions

If you wish to add a new model, you should follow different steps:

- create the model function

    ```{r, eval = FALSE}
    my_model <- function(...){
      ...
      ...
      ...
      class(OUT) = c("PPBstats", "fit_my_model")
      return(OUT)
    }
    ```

- create the _fit\_my\_model_ method for `check\_model()`

    ```{r, eval = FALSE}
    check_model.fit_my_model <- function(x){
      ...
      ...
      ...
      class(out) <- c("PPBstats", "check_my_model")
      return(OUT)
    }
    ```

- create the _check\_my\_model_ method for `mean\_comparisons()`

    ```{r, eval = FALSE}
    mean_comparisons.check_my_model <- function(x, ...){
      ...
      ...
      ...
      class(out) <- c("PPBstats", "mean_comparisons_my_model")
      return(out)
    }
    ```

- create the _mean\_comparisons\_my\_model_ method for `plot()`

    ```{r, eval = FALSE}
    plot.mean_comparisons_my_model <- function(x, y, ...){
      ...
      ...
      ...
    }
    ```



# How to contribute

1. If you want to interact with us, please file an __issue__ first to share your
intentions. We can help you plan the curse of action.

1. __Fork__ the git repository

1. Include your __contributions__ in your fork (preferably in a specific
branch)

1. Update the __documentation__ if necessary (help files, vignette)

1. Write unit __tests__ for your functions, if any

1. Pull the latest changes in the `master` branch of the upstream repository and
__rebase__ your work on top of them.

1. Run all tests and compile the vignette. Make sure everything __works__ as
expected.

1. Send a __Pull Request__ with a reference to the original _issue_




# Internal function structure


\begin{figure}
\begin{center}
\input{figures/PPBstats_secondary_functions.tex}
\end{center}
\caption{Internal functions used within main functions.}
\label{workflow_2}
\end{figure}

\begin{table}
\begin{tabular}{lp{.4\textwidth}}

\hline
\textbf{function name} & \textbf{description} \\ \hline

\texttt{get.env.info} & Get regional farms data and satellite farms data \\ \hline

\texttt{GxE\_build\_interaction\_matrix} & Compute interaction matrix \\ \hline

\texttt{check\_model\_model\_1} & Check if the \texttt{model\_1} model went well  \\ \hline

\texttt{check\_model\_model\_2} & Check if the \texttt{model\_2} model went well \\ \hline

\texttt{check\_model\_GxE} & Check if the GxE model went well \\ \hline

\texttt{parameter\_groups\_GxE} & Get matrix with variables in column and effect in row from \texttt{check\_model\_GxE} \\ \hline

\texttt{parameter\_groups\_model\_2} & Get matrix with variables in column and effect in row from \texttt{check\_model\_model\_2} \\ \hline

\texttt{comp.parameters} & Get parameter comparisons two by two or to a given threshold based on MCMC outputs \\ \hline

\texttt{get.significant.groups} & Get significant groups of differences for a set of parameters based on MCMC outputs \\ \hline

\texttt{get.at.least.X.groups} & Get the value of type one error needed to have X groups \\ \hline

\texttt{mean\_comparisons\_GxE} & Get mean comparisons from \texttt{check\_model\_GxE} \\ \hline

\texttt{mean\_comparisons\_model\_1} & Get mean comparisons from \texttt{check\_model\_model\_1} \\ \hline

\texttt{mean\_comparisons\_model\_2} &Get mean comparisons from \texttt{check\_model\_model\_2}  \\ \hline

\texttt{mean\_comparisons\_predict\_the\_past\_model\_2} & Get mean comparisons from \texttt{predict\_the\_past\_model\_2} \\ \hline

\texttt{ggplot\_parameter\_groups} & Get ggplot from \texttt{parameter\_groups} \\ \hline

\texttt{ggplot\_check\_model\_GxE} & Get ggplot from \texttt{check\_model\_GxE} \\ \hline

\texttt{ggplot\_check\_model\_model\_1} & Get ggplot from \texttt{check\_model\_model\_1} \\ \hline

\texttt{ggplot\_check\_model\_model\_2} & Get ggplot from \texttt{check\_model\_model\_2} \\ \hline

\texttt{ggplot\_cross\_validation\_model\_2} & Get ggplot from \texttt{cross\_validation\_model\_2} \\ \hline

\texttt{ggplot\_mean\_comparisons\_GxE} & Get ggplot from \texttt{mean\_comparisons\_GxE} \\ \hline

\texttt{ggplot\_mean\_comparisons\_model\_1} & Get ggplot from \texttt{mean\_comparisons\_model\_1} \\ \hline

\texttt{ggplot\_mean\_comparisons\_model\_2} & Get ggplot from \texttt{mean\_comparisons\_model\_2} \\ \hline

\texttt{ggplot\_mean\_comparisons\_predict\_the\_past\_model\_2} & Get ggplot from \texttt{mean\_comparisons\_predict\_the\_past\_model\_2} \\ \hline

\texttt{ggplot\_biplot\_GxE} & Get ggplot from \texttt{biplot\_GxE} \\ \hline

\texttt{ggplot\_discrimitiveness\_vs\_representativeness} & Get "discrimitiveness vs representativeness" ggplot from PCA on interaction matrix \\ \hline

\texttt{ggplot\_mean\_vs\_stability} & Get "mean vs stability" ggplot from PCA on interaction matrix \\ \hline

\texttt{ggplot\_which\_won\_where} & Get "which won where" ggplot from PCA on interaction matrix \\ \hline

\end{tabular}
\caption{Internal functions}
\label{function_descriptions_workflow_2}
\end{table}

\clearpage

Two additional funtions are also defined in the package:

\begin{tabular}{ll}
\texttt{common\_functions} & Some functions used in several functions of PPBstats \\
\texttt{gget.PPBstats.data} & Get PPBstats datas to run example of the vignette \\
\end{tabular}


