# Run the analysis

```{r}
library(PPBstats)
```


## Format data

```{r}
data("data_model_GxE")
data_model_GxE = format_data_PPBstats(data_model_GxE, type = "data_agro")
```


## Descriptive plots for each location

```{r}
vec_locations = levels(data_model_GxE$location)
list_hist_locations = lapply(
  vec_locations, function(x){
    p = plot(
      data_model_GxE, plot_type = "histogramm",
      vec_variables = c("y1")
      )
    p$y1
    }
  )
names(list_hist_locations) = vec_locations
```

Note that it is important that each element of the list refers to data of a given location in order to catch the right information in the next step when generating the report.

## GGE model for all locations

The function `workflow_gxe()` is coming from the book [here](https://priviere.github.io/PPBstats_book/family-2.html#ammi).

```{r}
workflow_gxe = function(x, gxe){
  out_gxe = model_GxE(data_model_GxE, variable = x, gxe_analysis = gxe)
  
  out_check_gxe = check_model(out_gxe)
  p_out_check_gxe = plot(out_check_gxe)
  
  out_mean_comparisons_gxe = mean_comparisons(out_check_gxe, p.adj = "bonferroni")
  p_out_mean_comparisons_gxe = plot(out_mean_comparisons_gxe)
  
  out_biplot_gxe = biplot_data(out_check_gxe)
  p_out_biplot_gxe = plot(out_biplot_gxe)

  out = list(
    "out_gxe" = out_gxe,
    "out_check_gxe" = out_check_gxe,
    "p_out_check_gxe" = p_out_check_gxe,
    "out_mean_comparisons_gxe" = out_mean_comparisons_gxe,
    "p_out_mean_comparisons_gxe" = p_out_mean_comparisons_gxe,
    "out_biplot_gxe" = out_biplot_gxe,
    "p_out_biplot_gxe" = p_out_biplot_gxe
  )
  
  return(out)
}


vec_variables = c("y1")
res_gge = lapply(vec_variables, workflow_gxe, "GGE")
names(res_gge) = vec_variables
```

## Create a list with all results

```{r}
res = list("hist_locations" = list_hist_locations, 
           "res_gge" = res_gge
           )
```

# Generate a report for each location

To generate the report, you need the `R` package `rmarkdown` installed.
In the following example, the output is `.html`. 
You can choose `.pdf` or `.docx`.
See `?rmarkdown::render` for more information.

Note that the template calls two objects:

- `params` which is list with parameter of the report
- `res` which is a list with all results coming from the analysis

The template used below can be download [here](https://github.com/priviere/PPBstats/blob/master/inst/return_report_web_site/_example_1_template.Rmd).


```{r, eval=FALSE}
library(rmarkdown)

vec_locations = names(res$hist_locations)
for (location in vec_locations){ # For each location, render a report
  params = list("title" = paste("Agronomic analyses for", location))
  rmarkdown::render("_example_1_template.Rmd", 
                    output_file =  paste("example_1_report_", location, ".html", sep = ""), 
                    output_dir = "./"
                    )
  }

```

The report generated can be visualized for 
[loc-1](example_1_report_loc-1.html), 
[loc-2](example_1_report_loc-2.html) and 
[loc-3](example_1_report_loc-3.html).




