
Here are the following step to install `PPBstats`:

1.**install `R`** on your computer.
Follow instruction at https://cloud.r-project.org/.
If your not familiar with `R`, it is advised to start with a tutorial (there are a lot on internet).

2. **install the `R` package `devtools`**
```{r, eval=FALSE}
install.packages("devtools")
```

3. **install `JAGS` software** on your computer in order to do bayesian analysis. 
See http://mcmc-jags.sourceforge.net/ for more details.

4. **download the current version of the package**: [download v0.24](https://github.com/priviere/PPBstats/archive/v0.24.zip)

5. **unzip the folder** on your local directory 

6. **install the package** thanks to the `install` funtion from the `devtool` package 
```{r, eval=FALSE}
devtools::install("PPBstats-0.24")
```
Note that you can also load the development version:
```{r, eval=FALSE}
devtools::install_github("priviere/PPBstats")
```
The evolution of the code regarding the development version are
[here](https://github.com/priviere/PPBstats/blob/master/NEWS.md).

7. **load the package** :
```{r,warning=FALSE,message=FALSE}
library(PPBstats)
```


